
---
title: "Country Indicators Analysis with R"
format: html
---

# Country Indicators Analysis with R

## Introduction

This document presents a visual analysis of country-level indicators such as population, GDP, GNI, and life expectancy using R.

### Scatter Plot: GDP per Capita vs. Life Expectancy

```{r}
library(ggplot2)
data <- read.csv('unicef_metadata.csv')
data_complete <-  subset(data, year == 2020)
#data_complete <- na.omit(data[c('GDP.per.capita..constant.2015.US..', 'Life.expectancy.at.birth..total..years.')])

ggplot(data_complete, aes(x=GDP.per.capita..constant.2015.US.., y=Life.expectancy.at.birth..total..years.)) +
  geom_point(aes(color=country), alpha=0.5) +
  labs(title="GDP per Capita vs. Life Expectancy", x="GDP per Capita (constant 2015 US$)", y="Life Expectancy at Birth (years)") +
  theme_minimal()
```

## Bar Chart: Population by Country (2020)

```{r}
data_2020 <- subset(data, year == 2020)

ggplot(data_2020[1:35,], aes(x=reorder(country, -Population..total), y=Population..total)) +
  geom_bar(stat="identity", fill="steelblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Population by Country in 2020", x="Country", y="Total Population")
```

## Map Visualization: GDP per Capita (2020)

```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
world <- ne_countries(scale = "medium", returnclass = "sf")
data_2020 <- subset(data, year == 2020)
data_2020$country_code <- tolower(data_2020$alpha_3_code)

world <- merge(world, data_2020, by.x = 'iso_a3', by.y = 'country_code', all.x = TRUE)

ggplot() +
  geom_sf(data = world, aes(fill = GDP.per.capita..constant.2015.US..)) +
  scale_fill_viridis_c() +
  labs(fill="GDP per Capita", title="GDP per Capita (2020) by Country") +
  theme_minimal()
```

## Findings

1. A clear positive correlation between GDP per capita and life expectancy indicates that wealthier countries tend to have higher life expectancies.
2. The population distribution in 2020 shows significant differences among countries, reflecting a wide range of demographic scales.
3. The map visualization of GDP per capita highlights economic disparities, with wealthier nations depicted in darker colors.

---

