
---
title: "Measles Vaccination Coverage Analysis with R"
format: html

---

# Measles Vaccination Coverage Analysis with R

## Introduction

This document presents a visual analysis of measles vaccination coverage across various countries.

## Bar Chart: Vaccination Coverage by Country in 2018

```{r}
#| warning: false

library(ggplot2)
data <- read.csv('Measles.csv')
data_2018 <- subset(data, time_period == 2018)

ggplot(data_2018, aes(x=reorder(country, -obs_value), y=obs_value, fill=country)) +
  geom_bar(stat="identity") +
  labs(title="Measles Vaccination Coverage by Country in 2018", x="Country", y="Coverage (%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Line Chart: Trends of Vaccination Over Time in Afghanistan

```{r}
#| warning: false

afghanistan_data <- subset(data, country == 'Afghanistan')

ggplot(afghanistan_data, aes(x=time_period, y=obs_value, group=1)) +
  geom_line() +
  geom_point() +
  labs(title="Trends of Measles Vaccination Over Time in Afghanistan", x="Year", y="Coverage (%)")
```

## Map Visualization: Global Distribution of Measles Vaccination Coverage (2018)

```{r}
#| warning: false

library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
world <- ne_countries(scale = "medium", returnclass = "sf")
data_2018$country_code <- tolower(data_2018$alpha_3_code)

world <- merge(world, data_2018, by.x = 'iso_a3', by.y = 'alpha_3_code', all.x = TRUE)

ggplot() +
  geom_sf(data = world, aes(fill = obs_value)) +
  scale_fill_viridis_c() +
  labs(fill="Vaccination Coverage (%)", title="Global Distribution of Measles Vaccination Coverage (2018)")
```

## Findings

1. Measles vaccination coverage in 2018 shows significant global disparities with some countries achieving high coverage rates while others lag significantly.
2. The trend in Afghanistan indicates fluctuating measles vaccination rates over the years, pointing to challenges in healthcare delivery and policy implementation.
3. The map visualization highlights countries with insufficient coverage, suggesting areas where international health interventions could be crucial.

---

```{r}
#| warning: false
 
library(dplyr)
library(tidyr)

data_c <- read.csv('Children under 5 Mortality_v1.csv')

summary_stats <- data_c %>%
  group_by(ParentLocation, Period) %>%
  summarize(Mean_Child_Mortality = mean(child.mortality, na.rm = TRUE),
            Median_Child_Mortality = median(child.mortality, na.rm = TRUE),
            Min_Child_Mortality = min(child.mortality, na.rm = TRUE),
            Max_Child_Mortality = max(child.mortality, na.rm = TRUE))


```
## Line Chart: Plot for Child Mortality

```{r}
#| warning: false

#time_series_plot <- 
  ggplot(data_c, aes(x = Period, y = child.mortality, group = SpatialDimValueCode, color = SpatialDimValueCode)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Time Series of Child Mortality",
       x = "Year",
       y = "Child Mortality Rate",
       color = "Country")


```
## Line Chart: Plot for Child Mortality

```{r}
#| warning: false

#time_series_plot <- 
 ggplot(data_c, aes(x = ParentLocation, y = child.mortality, fill = ParentLocation)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Child Mortality Rates by Location",
       x = "Location",
       y = "Child Mortality Rate")


```
